---
title: "Cod landings project"
author: "Kristin, Delphine, Reese"
date: "2023-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r upload-packages}
#install.packages("tidyverse")
library(tidyverse)
#install.packages("devtools")
library(devtools)
#devtools::install_github("rstudio-education/dsbox", force = TRUE)
library(dsbox)
#install.packages("ggridges")
library(ggridges)
#install.packages("gganimate")
library(gganimate)
#install.packages("janitor")
library(janitor)
#install.packages("stringr")
library(stringr) ## String manipulation
#install.packages("plotly")
library(plotly)
#install.packages("usmap")
library(usmap)
library(ggplot2)
#install.packages("devtools")
#devtools::install_github("UrbanInstitute/urbnmapr")
library(urbnmapr)
#install.packages("leaflet")
library(leaflet) ## For leaflet interactive maps
#install.packages("sf")
library(sf) ## For spatial data
#install.packages("RColorBrewer")
library(RColorBrewer) ## For colour palettes
#install.packages("htmltools")
library(htmltools) ## For html
#install.packages("leafsync")
library(leafsync) ## For placing plots side by side
#install.packages("kableExtra")
library(kableExtra) ## Table output
#install.packages("stringr")
library(stringr) ## String manipulation
library(gganimate)
library(transformr)#install.packages("ggimage")
library(ggimage)
#install.packages("naniar")
library(naniar)
#install.packages("visdat")
library(visdat)
```

```{r us_maps-shapefile}

us_maps <- us_map(region = "states")

states_sf <- get_urbn_map(map = "states", sf = TRUE)


```



```{r modify-column-names}
cod_landings <- read_csv("data/cod-landings.csv")  
cod_landings <- cod_landings 
names(cod_landings)[names(cod_landings) == "NMFS Name"] <- "Common_Name"
names(cod_landings)[names(cod_landings) == "Metric Tons"] <- "Metric_Tons"
names(cod_landings)[names(cod_landings) == "Dollars"] <- "US_Dollars"
names(cod_landings)[names(cod_landings) == "Scientific Name"] <- "Scientific_Name"
names(cod_landings)[names(cod_landings) == "Collection"] <- "Fishing_Types"
names(cod_landings)[names(cod_landings) == "Tsn"] <- "Taxonomic_Serial_Number"
names(cod_landings)[names(cod_landings) == "Source"] <- "Fishing_Companies"
colnames(cod_landings)

```
```{r glimpe-dataset}
glimpse(cod_landings)
```

```{r change-wording-dataset}
cod_landings$Common_Name <- str_replace(cod_landings$Common_Name,"COD, PACIFIC", "Pacific_Cod")

cod_landings$Common_Name <- str_replace(cod_landings$Common_Name,"COD, ATLANTIC", "Atlantic_Cod")

cod_landings$Scientific_Name <- str_replace(cod_landings$Scientific_Name, "Gadus macrocephalus", "Gadus_macrocephalus")

cod_landings$Scientific_Name <- str_replace(cod_landings$Scientific_Name, "Gadus morhua", "Gadus_morhua")
cod_landings
```

Q1 - How do the weights of recorded recreational and commercial Cod landings of the Atlantic and Pacific compare to one another between the years 1950 and 2021?
Plan Q1 - Summarize the data through a ridgeline density plot of weight of landings, colored by “Fishing_Type” and faceted by “Common_Name”. Also include an animated plot of the same type to better show the fluctuation in landings every 10 years. 

```{r filter-out-arctic-and-toothed-cod}
cod_landings <- cod_landings %>% 
  filter(Common_Name != "COD, TOOTHED" ) %>% 
  filter(Common_Name != "COD, ARCTIC" )
```

```{r mutate-total-metric-tons-column}
cod_landings_total <- cod_landings %>%
  group_by(Year) %>%
   mutate(Total_Metric_Tons = sum(Metric_Tons, na.rm = TRUE)) %>%
  arrange(desc(Year))
view(cod_landings_total)

```

```{r mutate-total-pounds-column}
cod_landings_total <- cod_landings %>%
  group_by(Year) %>%
   mutate(Total_Pounds = sum(Pounds, na.rm = TRUE)) %>%
  arrange(desc(Year))
```

```{r count-total-ponds-choose-sclae-axis-graph}
cod_landings_total %>% 
  group_by(Total_Pounds) %>% 
count(Total_Pounds) %>% 
  arrange(desc(Total_Pounds))
```

```{r cod-image}
cod_image <- "pacific-cod.webp"
cod_landings_total %>%
  mutate(
    image = case_when(Year %in% c(1950:2021) ~ cod_image,
                      TRUE ~ cod_image)) 
```

```{r cod-image-graph-with-path, fig.alt="Faceted animated plot using an image of a cod leaving a trace and forming an area to show the variation in cod landings by millions of pounds from 1950 to 2021 for the Atlantic and the Pacific cod. The values for Pacific cod landings are in general smaller than for the Atlantic cod. The Pacific cod landings peak in a few years around the ranges of 4000 millions of tons in late 1990's early 2000 and for a few years in the 2010's. Altlantic cods in these same years are larger and closer to two or three times the values for Pacific cod.The Atlantic cod peak mid 1990 and early 2010's".}
#should we include more precise dates and the actual axis and title to the alt-text?
#should we remove the colors?
#should we clean this code?

fish_coords <- cod_landings_total %>%
  group_by(Year, Common_Name) %>%
  summarize(Total_Pounds = sum(Total_Pounds, na.rm = T))


Landings_Overview_Graph <- fish_coords %>% 
  ggplot(mapping = aes(
         x = Year,
         y = Total_Pounds/1000000)) +
  geom_col(data = cod_landings_total, aes(fill = Common_Name))+
  geom_path(data = fish_coords) +
#  geom_point(data = fish_coords) +
  geom_image(data = fish_coords, aes(image = cod_image), size = 0.1, alpha = 0.3) + 
  scale_color_manual(values = fill) +
    scale_x_continuous(breaks = seq(1950, 2021, 10))+
    #scale_y_continuous(breaks = seq(40000000,730000000, 500000000))+
  facet_wrap (~ Common_Name, nrow=2, scales = "free_x") +
  scale_fill_viridis_d() +
   labs(title = "US Pacific and Atlantic Cod Landings From 1950 to 2021",
        subtitle = "In Millions of Pounds",
        x = "Years",
        y = "Millions of Pounds",
        fill = "Fish Species' Common Name") +
   transition_reveal(Year)

#ggsave("Landings_Overview_Graph.png",Landings_Overview_Graph, width = 6, height = 8)

Landings_Overview_Graph_gif <- image_join(lapply(frames, function(Landings_Overview_Graph) {
  print(Landings_Overview_Graph)
  as.raster(dev.capture())}))

image_write(gif, "Landings_Overview_Graph_gif")
```


```{r Q1-interactive-landings-by-fishing-types, fig.height=8, fig.width=6, fig.alt= "Interactive graph of recorded cod landings as a stacked bar chart by commercial and recreational fishing type faceted for the Atlantic and Pacific Cod. There is no record of recreational data for any species before 1980. For the Atlantic cod, there is an overall variation of landings of recreational and commercial fishing over the years. Starting around 2016, the amount of commercial landings decreases at a faster rate than those of commercial fishing. Before 2016, there is mostly a mojority of commercial fishing, with a few years where recreational fishing has higher landings in the early and late 2000's. For the Pacific cod, there is overal few few data of commercial recorded landings before mid 1990's. From 2000 to 2021, there is a closer ratio of recorded commercial and recreational landings".}

#change the total pounds column in dataset so that it is /1000000 too so that when you click on the interactive chart the info you see is also more accessible (like on the y-axis)
#figure out how to add labels
#add "accessible" colors?
#add information about the graph text
cod_landings_total <- cod_landings_total %>% 
  mutate(Million_Pounds = Total_Pounds/1000000)

Landings_By_Type_Plotly <- cod_landings_total %>% 
  group_by(Common_Name) %>%
  ggplot(mapping = aes(
         x = Year,
         y = Million_Pounds,
         fill = Fishing_Types)) +
    geom_col(alpha=0.5) +
   # geom_line() +
 #   geom_vline(xintercept = 1992, color = "red" ) +
  #  geom_text(aes(v_line, label = "1992 Cod Moratorium")) +
    scale_x_continuous(breaks = seq(1950, 2021, 10))+
    labs(title = "Atlantic and Pacific Cod Landings By Fishing Types",
         subtitle = "From 1950 to 2021",
         x = "Years",
         y = "Millions of Pounds",
         fill = "Fishing Type")+
  facet_wrap( ~ Common_Name, nrow = 2, scales = "free_x")

Landings_By_Type_Plotly <- ggplotly(Landings_By_Type_Plotly) 
Landings_By_Type_Plotly

#plotly::saveWidget(Landings_By_Type_Plotly, file = "Landings_By_Type_Plotly.html")

# if want to choose color --> geom_area(fill="#69b3a2", alpha=0.5) +
 # geom_line(color="#69b3a2") 


```


QUESTION 2

```{r rename} {col-capitailzation}
states_sf <- states_sf %>% 
  mutate(state_name = tolower(state_name))

states_sf <- states_sf %>%
  rename("State" = "state_name")
 

cod_landings <- cod_landings %>%
  mutate(State = tolower(State))
```

```{r commercial-profit}
cod_landings_commercial <- cod_landings %>%
  na.omit(US_Dollars) %>%
  rename(Commercial_Profit = US_Dollars)
```

```{r transform-sf}
states_sf <- st_transform(states_sf, "+init=epsg:4326")
st_crs(states_sf)
```

```{r join-shapefile} {commercial}
cod_landings_commercial_total <- cod_landings_commercial %>% 
  full_join(states_sf, by = "State")

cod_landings_commercial_total <- st_sf(cod_landings_commercial_total)
```

```{r join-shaefile} {commercial-recreational}
cod_landings_sf <- cod_landings %>%
full_join(states_sf, by = "State")

cod_landings_sf <- st_sf(cod_landings_sf)
```


```{r US_profit-through-years} 
 cod_landings_commercial_total %>%
  ggplot(mapping = aes(
         x = Year,
         y = Commercial_Profit/1000000000)) +
    geom_col(alpha=0.5) 

#ggplotly( cod_landings_commercial_total)
```


```{r, fig.alt= "The facet maps of the US repersents the monitary value of Pacific and Atlantic Cod in relation to location and year. The years displayed are 1950, 1980, 1999, 2008, 2018, and 2021. Overall monitary benifits from Cod landings incleased over time until 2008; there was a spike and then decline. Alaska gains the most monitary value from cod landings compared to other states."}   {US-commercial-ecolomy-map} 

cod_landings_commercial_total %>%
  filter(Year %in% c(1950, 1980, 1999, 2008, 2018, 2021)) %>% 
  ggplot() +
  geom_sf(aes(fill = Commercial_Profit/100000000))+  #figure out expand
 facet_wrap(. ~ Year)+
  #transition_time(time = Year)
  labs(title = " US Pacific and Atlantic Cod Landings Regional Economy From 1950 to 2021", 
       subtitle = "In Trillions of Dollars",
       fill= "Profit (trillions)",
       x = "Longitude", 
       y = "Latitude")

#ggplotly(cod_landings_commercial_total)

#would be intersting to know why there's such a large decrease from 2008-2009
```
```{r Landings-by-state} {Recreational-commercial}
cod_landings_sf %>%
ggplot(mapping = aes(
         x = Year,
         y = Pounds/10000000,
       fill = State )) +
  geom_col(alpha=0.4)+
 # scale_fill_viridis_d()
   labs(title = " US Pacific and Atlantic Cod Landings From 1950 to 2021", 
       subtitle = "By State",
       fill= "State",
       x = "Year", 
       y = "Billions of Dollars")

#ggplotly(cod_landings_sf)
```

```{r Economy-by-state-over-years} {Commercial-only}
cod_landings_commercial_total %>%
ggplot(mapping = aes(
         x = Year,
         y = Commercial_Profit/1000000000,
       fill = State )) +
  geom_col(alpha=0.4)+
 # scale_fill_viridis_d()
   labs(title = " US Pacific and Atlantic Cod Landings Economy From 1950 to 2021", 
       subtitle = "By State",
       fill= "State",
       x = "Year", 
       y = "billions of Dollars")

#ggplotly(cod_landings_commercial_total)
```


```{r Alaska-Washington-landing} {Recreational-commercial}
cod_landings_sf %>%
  filter(State %in% c("alaska", "washington")) %>%
  ggplot()+
  geom_sf(aes(fill = Pounds/1000000))+
  facet_grid(. ~ State)+
  transition_time(time = Year)
  labs(title = "Alaska and Washington's Pacific Cod Landings From 1950 to 2021", 
       subtitle = "In Millions of Pounds",
       fill= "Millions of Pounds",
       x = "Longitude", 
       y = "Latitude")
```

```{r Alaska-Washington-economy} {Commercial-only}
cod_landings_commercial_total %>%
  filter(State %in% c("alaska", "washington")) %>%
  ggplot()+
  geom_sf(aes(fill = Commercial_Profit/100000000))+
  facet_grid(. ~ State)+
  transition_time(time = Year)
  labs(title = "Alaska and Washington's Commercial Pacific Cod Landings Economy From 1950 to 2021", 
       subtitle = "In Billions of Dollars",
       fill= "Billions of Dollars",
       x = "Longitude", 
       y = "Latitude")

```

```{r Maine-North-Carolina-landings} {Recreational-commercial}
cod_landings_sf %>%
  filter(State %in% c("maine", "north carolina")) %>%
  ggplot()+
  geom_sf(aes(fill = Pounds/1000000))+
  facet_grid(. ~ State)+
  transition_time(time = Year)
  labs(title = "Maine and North Carolina's Atlantic Cod Landings From 1950 to 2021", 
       subtitle = "In Millions of Pounds",
       fill= "Millions of Pounds",
       x = "Longitude", 
       y = "Latitude")
```

```{r Maine-North-Carolina-economy} {Commercial-only}
cod_landings_commercial_total %>%
  filter(State %in% c("maine", "north carolina")) %>%
  ggplot()+
  geom_sf(aes(fill = Commercial_Profit/100000000))+
  facet_grid(. ~ State)+
  transition_time(time = Year)
  labs(title = "Maine and North Carolina's Commercial Atlantic Cod Landings Economy From 1950 to 2021", 
       subtitle = "In Billion of Dollars",
       fill= "Billions of Dollars",
       x = "Longitude", 
       y = "Latitude")

```


```{r missing-data}
cod_landings %>%
  ggplot(aes(x = Year, y = Pounds, color= Fishing_Types))+
  geom_point()+
  geom_miss_point()
```

