---
title: "Codlandingsproject"
author: "Kristin, Delphine, Reese"
date: "2023-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r upload-packages}
#install.packages("tidyverse")
library(tidyverse)
#install.packages("devtools")
library(devtools)
devtools::install_github("rstudio-education/dsbox", force = TRUE)
library(dsbox)
#install.packages("ggridges")
library(ggridges)
#install.packages("gganimate")
library(gganimate)
```

```{r modify-column-names}
cod_landings <- read_csv("data/cod-landings.csv")
cod_landings <- cod_landings 
  colnames(cod_landings)[3] = "Common_Name"
  colnames(cod_landings)[5] = "Metric_Tons"
  colnames(cod_landings)[6] = "US_Dollars"
  colnames(cod_landings)[7] = "Fishing_Types"
  colnames(cod_landings)[8] = "Scientific_Name"
  colnames(cod_landings)[9] = "Taxonomic_Serial_Number"
  colnames(cod_landings)[10] = "Fishing_Companies"
```
```{r glimpe-dataset}
glimpse(cod_landings)
```
Q1 - How do the weights of recorded recreational and commercial Cod landings of the Atlantic and Pacific compare to one another between the years 1950 and 2021?
Plan Q1 - Summarize the data through a ridgeline density plot of weight of landings, colored by “Fishing_Type” and faceted by “Common_Name”. Also include an animated plot of the same type to better show the fluctuation in landings every 10 years. 
```{r Q1-col-trial}
#Maybe a stacked area chart would be better, review the types of vizualisation then create a static one and after try the animated 
ggplot(data = cod_landings,
       mapping = aes(
         x = Year,
         y = Pounds,
         fill = Common_Name))+
  geom_col()+
  facet_wrap( ~ Fishing_Types)
```
```{r Q1-Density-trial}
#the mutating is not working
#Density doesn't represent the pounds... find another plot
cod_landings %>% 
  mutate(Atlantic = if_else(Common_Name == "COD, ATLANTIC", 
                           "Atlantic", 
                           "Pacific" )) %>% 
  ggplot(mapping = aes(x = Year, 
                       fill = Fishing_Types, 
                       color = severity ))+
  geom_density(alpha = 0.5, 
               color = "black")

#change graduation on the axis
#Add labs
#Change color
#Make it move
```


